CSDA in 2D slices of the hippocampus \emph{done right}

### 2015-06-05 Reportes/CSDADoneRight01.tex
cambios pecuecos

This expressions may be regardes as inescesarily long, as the 
$\cond$ tensor, being a real symmetric tensor, can be represented
in local coordinates which diagonalize it. Those coordinates have
physical meaning, as they are the resolving directions in which
conductivity behaves differently and denote some underlying structure
in the studied material. The usual assumption is that 
the conductivity tensor is isotropic but not homogeneous, and it 
changes along the apical direction, and not along the 
differentiated strata (e.g. stratum pyramidale, stratum granolosum).
Let us suppose that this is approximately correct, and that very close
to the stratum pyramidale we represent $\cond$ into the coordinates
defined by its eigenvectors (see fig. \ref{esquemas01}).
 We shall call them $(t,a)$ coordinates
and asume that the first is approximately tangent to the
strata and the other to the apical direction. The $\cond$ tensor would 
have the expression in such coordinates
\begin{equation}
\tilde{\cond}=
\begin{pmatrix}
\sigma_t & 0 \\
0 & \sigma_a
\end{pmatrix}
\end{equation} 
Then, the equation \ref{Imcart} becomes
\begin{equation}\label{Imapic}
\begin{split}
-I_m= & (\partial_t \sigma_{t}) (\partial_t \phi) +
(\partial_a \sigma_{a}) (\partial_a \phi)  \\
& + \sigma_{a} \partial_a^2 \phi + \sigma_{t} \partial_t^2 \phi, 
\end{split}
\end{equation}
as $(t,a)$ are also local ortogonal coordinates. 
One may wonder why we are explicitly stating the non-diagonal 
version of the equations. The data as it is delivered by
the measuring apparatus is on a strictly cartessian grid, and
we have to perform the numerical derivatives over this data, so
we have to know how these assumptions appear in both sets
of coordinates. Besides, the \emph{tangent-apical} coordinates
are just a teoretical tool, a real implementation of such a coordinate
set would be extremely complicated and prone to numerical errors.

If the medium where the ions diffuse is liquid, and the structures
therein are randomly distributed, the assumption of isotropy holds, 
$\sigma_t=\sigma_a=\sigma$. Then, as we stated, there is no
change along the strata, but there is change along the appical
direction:
\begin{align}
\partial_t \sigma & =0 \\
\partial_a \sigma & \neq 0. 
\end{align}
We are then left with
\begin{equation}
I_m=-\partial_t \sigma \partial_t \phi - \sigma \nabla^2 \phi.
\end{equation}
Here we must point out that the Laplacian is an invariant operator,
so the last summand of the equation could be in principle be implemented
in \emph{any} convenient set of coordinates, but in numerical aplications,
the \emph{Discrete Laplacian Operator} is \emph{not} invariant (further
discution in the following section). 
The term $\partial_t \sigma \partial_t \phi$ is the hindrance to 
ionic displacement due to dense layers of soma and dendrites near the
CA structure. Most autors have neglected its efects, and in one
dimensional analysis the structure of sources and sinks is not affected
by this. Now, as the aquisition of data in two dimensions becomes
finer, this term could lead to inexactitude in the location of the
sinks and sources, depending on the magnitude of the
derivative of the conductance. Locally both $(x,y)$ and
$(t,a)$ are orthonormal systems, so we can change from one to another 
with a rotation. If $theta$ is the angle between the $x$ vector
and $t$, the partial derivatives follow the next set of relations:
\begin{align}
\partial_t \sigma &= \cos\theta \partial_x \sigma +\sin\theta \partial_y \sigma=0\\
\partial_a \sigma &= -\sin\theta \partial_x \sigma +\cos\theta \partial_y \sigma
\end{align}

>>>>

This expressions may be regardes as inescesarily long, as the 
$\cond$ tensor, being a real symmetric tensor, can be represented
in local coordinates which diagonalize it. Those coordinates have
physical meaning, as they are the resolving directions in which
conductivity behaves differently and denote some underlying structure
in the studied material. The usual assumption is that 
the conductivity tensor is isotropic but not homogeneous, and it 
changes along the apical direction, and not along the 
differentiated strata (e.g. stratum pyramidale, stratum granolosum).
Let us suppose that this is approximately correct, and that very close
to the stratum pyramidale we represent $\cond$ into the coordinates
defined by its eigenvectors (see fig. \ref{esquemas01}).
 We shall call them $(t,a)$ coordinates
and asume that the first is approximately tangent to the
strata and the other to the apical direction. The $\cond$ tensor would 
have the expression in such coordinates
\begin{equation}
\tilde{\cond}=
\begin{pmatrix}
\sigma_t & 0 \\
0 & \sigma_a
\end{pmatrix}
\end{equation} 
Then, the equation \ref{Imcart} becomes
\begin{equation}\label{Imapic}
\begin{split}
-I_m= & (\partial_t \sigma_{t}) (\partial_t \phi) +
(\partial_a \sigma_{a}) (\partial_a \phi)  \\
& + \sigma_{a} \partial_a^2 \phi + \sigma_{t} \partial_t^2 \phi, 
\end{split}
\end{equation}
as $(t,a)$ are also local ortogonal coordinates. 
One may wonder why we are explicitly stating the non-diagonal 
version of the equations. The data as it is delivered by
the measuring apparatus is on a strictly cartessian grid, and
we have to perform the numerical derivatives over this data, so
we have to know how these assumptions appear in both sets
of coordinates. Besides, the \emph{tangent-apical} coordinates
are just a teoretical tool, a numerical implementation of such a coordinate
set would be  complicated and prone to numerical errors.

If the medium where the ions diffuse is liquid, and the structures
therein are randomly distributed, the assumption of isotropy may hold 
on average in sufficiently big scales \cite{Bedard11}. 
In order to not make the notation more cumbersome, let us aggree
that we are dealing with averages over volumes of the aproximate size
of the electrodes. Then, 
$\sigma_t=\sigma_a=\sigma$. As we stated, there is no
change along the strata, but there is change along the appical
direction:
\begin{align}
\partial_t \sigma & =0 \\
\partial_a \sigma & \neq 0. 
\end{align}
We are then left with
\begin{equation}
I_m=-\partial_t \sigma \partial_t \phi - \sigma \nabla^2 \phi.
\end{equation}
Here we must point out that the Laplacian is an invariant operator,
so the last summand of the equation could be in principle be implemented
in \emph{any} convenient set of coordinates, but in numerical aplications,
the \emph{Discrete Laplacian Operator} is \emph{not} invariant (further
discution in the following section). 
The term $\partial_t \sigma \partial_t \phi$ is the hindrance to 
ionic displacement due to dense layers of soma and dendrites near the
CA structure. Most autors have neglected its efects, and in one
dimensional analysis the structure of sources and sinks is not affected
by this. Now, as the aquisition of data in two dimensions becomes
finer, this term could lead to inexactitude in the location of the
sinks and sources, depending on the magnitude of the
derivative of the conductance. Locally both $(x,y)$ and
$(t,a)$ are orthonormal systems, so we can change from one to another 
with a rotation. If $theta$ is the angle between the $x$ vector
and $t$, the partial derivatives follow the next set of relations:
\begin{align}
\partial_t \sigma &= \cos\theta \partial_x \sigma +\sin\theta \partial_y \sigma=0\\
\partial_a \sigma &= -\sin\theta \partial_x \sigma +\cos\theta \partial_y \sigma
\end{align}
